Bird beak's elaboration and exploration: data preparation

### 2021-08-27 Analysis/01-Data_preparation.Rmd
Update 01-Data_preparation.Rmd

 * The clade data

@@

#@param table the clade list data frame
#@param level the ID of the column to use as a level
#@param threshold drop levels with less than the threshold
make.level.list <- function(table, level, threshold = 20) {
  ## Select the level
  table_level <- table[, level]
  ## Remove the empties
  table_level <- as.factor(as.character(table_level[-which(table_level == "")]))
  ## Select the levels below the threshold
  remove_levels <- which(table(table_level) < threshold)
  ## Remove the levels
  while(length(remove_levels) > 0) {
    table_level <- as.factor(as.character(table_level[-which(table_level == names(remove_levels)[1])]))
    remove_levels <- remove_levels[-1]
  }
  ## Convert into a list
  factors <- levels(table_level)
  levels_list <- list()
  while(length(factors) > 0) {
    levels_list[[length(levels_list) + 1]] <- which(table_level == factors[1])
    factors <- factors[-1]
  }
  ## Name the elements correctly
  names(levels_list) <- levels(table_level)
  return(levels_list)
}

## Reading the clades list
groups_level1 <- make.level.list(clades_list, level = 1, threshold = 20)
groups_level2 <- make.level.list(clades_list, level = 2, threshold = 20)
all_groups <- c(groups_level1, groups_level2)
```

>>>>

 * The clade data
We used the different clade levels (the 11 magnificent 7 + the families). For each level/group, we removed any groups with less than 20 species.

@@

#@param table the clade list data frame
#@param level the ID of the column to use as a level
#@param threshold drop levels with less than the threshold
make.level.list <- function(table, level, threshold = 20) {
  ## Select the level
  table_level <- table[, level]
  ## Find the empty level
  empty_level <- which(levels(table_level) == "")
  if(length(empty_level) != 0) {
      ## Remove the empties
      level_list <- as.factor(as.character(table_level[-which(table_level == "")]))
  }
  empty_level <- ""
  ## Select the levels below the threshold
  remove_levels <- which(table(level_list) < threshold)
  ## Remove the levels
  while(length(remove_levels) > 0) {
    level_list <- as.factor(as.character(level_list[-which(level_list == names(remove_levels)[1])]))
    table_level[which(table_level == names(remove_levels)[1])] <- empty_level
    table_level <- as.factor(as.character(table_level))
    remove_levels <- remove_levels[-1]
  }
  ## Convert into a list
  factors <- levels(level_list)
  list_levels <- list()
  while(length(factors) > 0) {
    list_levels[[length(list_levels) + 1]] <- which(level_list == factors[1])
    factors <- factors[-1]
  }
  ## Name the elements correctly
  names(list_levels) <- levels(level_list)

  ## Also returns the new column
  return(list(col = table_level, list = list_levels))
}

## Reading the clades list
groups_level1 <- make.level.list(clades_list, level = 1, threshold = 20)
groups_level2 <- make.level.list(clades_list, level = 2, threshold = 20)
all_groups <- c(groups_level1$list, groups_level2$list)
```
